---
title: "First Steps with Rcpp"
---

To test our compiler and make sure that Rcpp is installed, execute the following to run and compile a short Rcpp snippet

```{r test_compiler, code=readLines(here('R', 'code', 'compiler_test.R'))}
```

The above code, although seemingly short, is in fact doing a whole lot in the background. This call to `evalCpp` is taking the passed in string, constructing the necessary boilerplate code for a minimal C++ program, then compiling and executing the resulting program. 

It's also possible to compile a string into a function that is then accessible to you in R

```{r inline_func, code=readLines(here('R', 'code', 'first_cpp_func.R'))}
```

One of the primary reasons to use Rcpp is for gains in speed as we're working at a lower abstraction level. Let's see how this cpp function compares to the R version

```{r inline_benchmark, code=readLines(here('R', 'code', 'inline_benchmark.R'))}
```

The C++ version of `is_odd` is about twice as fast as the R version.

It's easy to see, though, that this is inline usage of strings is not a scalable way of writing complex C++ code to be called from R. Luckily, Rcpp works very well using the original C++ file format `.cpp`

